---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# FARS

FARS has been developed in a context of the Coursera's course "[Building an R Package](https://www.coursera.org/learn/r-packages/home/welcome)". 
The package provides a few very basic functions, which work on top of embedded in the package sample FARS data (for more information about FARS data please see [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)).

## Installation

You can install FARS from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Valentin-Konoshenko/FARS")
```

## Usage

It probably makes sense to start analysis FARS data from exprloring the big picture. It is where 
`fars_summarize_years` comes into play.
```{r}
## exploring summarized data

FARS::fars_summarize_years(2013:2015)
```
You can focus on a specific state and a year and plot the accidents on a map using `fars_map_state`:
```{r}
## plotting the accidents on a map

FARS::fars_map_state(6, 2015)
```
The `fars_read` function will be useful if you want to work with raw data directly. For instanse let's find out the least and the most dangerous hours  
```{r}
## the least and the most dangerous hours

library(dplyr)

data_file <- FARS::make_filename(2015)
summarized_data <- FARS::fars_read(data_file) %>%
  filter(HOUR != 99) %>%
  group_by(HOUR) %>% 
  summarise(number_of_cases = n()) %>%
  arrange(number_of_cases)
d <- summarized_data[c(1, nrow(summarized_data)), ]
cat("\nThe least dengures hour:", unlist(d[1, "HOUR"]), 
    "\nThe most dengures hour: ", unlist(d[2, "HOUR"]))
```
