---
title: "Overview"
author: "Valentin Konoshenko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

FARS has been developed in a context of the Coursera's course "[Building an R Package](https://www.coursera.org/learn/r-packages/home/welcome)". 
The package provides a few very basic functions, which work on top of embedded in the package sample FARS data (for more information about FARS data please see [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)).

## Usage

### Exploring summarized data

It probably makes sense to start analysis FARS data from exprloring the big picture. It is where 
`fars_summarize_years` comes into play.
```{r}
FARS::fars_summarize_years(2013:2015)
```
### Plotting the accidents on a map

You can focus on a specific state and a year and plot the accidents on a map using `fars_map_state`:
```{r, fig.show='hold'}
FARS::fars_map_state(6, 2015)
```

### Exploring raw data

The `fars_read` function will be useful if you want to work with raw data directly. For instanse let's find out the least and the most dangerous hours  
```{r}
## the least and the most dangerous hours

library(knitr)
library(dplyr)

data_file <- FARS::make_filename(2015)
summarized_data <- FARS::fars_read(data_file) %>%
  filter(HOUR != 99) %>%
  group_by(HOUR) %>% 
  summarise(number_of_accidents = n()) %>%
  arrange(number_of_accidents)
d <- rbind(
  head(summarized_data, 1),
  tail(summarized_data, 1))
knitr::kable(d)

cat("\nThe least dangerous hour:", unlist(d[1, "HOUR"]), 
    "\nThe most dangerous hour: ", unlist(d[2, "HOUR"]))
```